\documentclass{article}

% packages for math
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}

\usepackage{enumitem}

% package for including images
\usepackage{graphicx}

% TAKEN FROM OVERLEAF DOCUMENTATION
% https://www.overleaf.com/learn/latex/Code_listing
\usepackage{listings}
\lstset{language=Python}
\usepackage{xcolor}
\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}
\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\ttfamily\footnotesize,
    breakatwhitespace=false,
    breaklines=true,
    captionpos=b,
    keepspaces=true,
    numbers=left,
    numbersep=5pt,
    showspaces=false,
    showstringspaces=false,
    showtabs=false,
    tabsize=2
}
\lstset{style=mystyle}

\newcommand{\vv}[1]{\mathbf{#1}}
\newcommand{\vspan}{\mathsf{span}}
\newcommand{\ran}{\mathsf{ran}}
\newcommand{\cod}{\mathsf{cod}}
\newcommand{\R}{\mathbb R}

% environment for solutions
\theoremstyle{remark}
\newtheorem*{solution}{Solution}

% capital letters for problem parts
\renewcommand{\theenumi}{\Alph{enumi}}

% no page numbers
\pagenumbering{gobble}

% UNCOMMENT IF YOU DON'T WANT PROBLEMS ON INDIVIDUAL PAGES
% \renewcommand{\pagebreak}{}

\title{
  Week 5 Discussion
}
\author{CAS CS 132: Geometric Algorithms}
\date{October 2, 2023}

\begin{document}
\maketitle

\noindent During discussion sections, we will go over three problems.
\begin{itemize}
\item The first will be a warm-up question, to help you verify your understanding of the material.
\item The second will be a solution to a problem on the assignment of the previous week.
\item The third will be a problem similar to one on the assignment of the following week.
\end{itemize}
The remainder of the time will be dedicated to open Q\&A.
\pagebreak

\section{Definitions}

Let
$
A =
\begin{bmatrix}
  \mathbf a_1 & \mathbf a_2 & \dots & \mathbf a_n
\end{bmatrix}
$
be a $m \times n$ matrix (in particular, each column $\mathbf a_i$ is a vector in $\mathbb R^m$), let $\mathbf b$ be an arbitrary nonzero vector in $\mathbb R^m$ and let $T$ be the linear transformation implemented by $A$.
Consider the following list of statements.
\begin{enumerate}[label=\arabic*.]
\item The equation $A \mathbf x = \mathbf b$ is consistent.
\item The equation $A \mathbf x = \mathbf c$ is consistent for any choice of $\mathbf c$.
\item $A \mathbf x = \mathbf b$ has a unique solution.
\item $A \mathbf x = \mathbf 0$ has a unique solution.
\item $A \mathbf x = \mathbf c$ has a unique solution for any choice of $\mathbf c$.
\item $\mathbf b \in \vspan\{\mathbf a_1, \mathbf a_2, \dots, \mathbf a_n\}$.
\item $\mathbf c \in \vspan\{\mathsf a_1, \mathbf a_2, \dots, \mathbf a_n\}$ for any choice of $\mathbf c$.
\item $\vv a_n \in \vspan\{\vv a_1, \vv a_2, \dots, \vv a_{n - 1}\}$.
\item $A$ has a pivot in every column.
\item $A$ has a pivot in every row.
\item $\begin{bmatrix} \vv a_1 & \vv a_2 & \dots & \vv a_n & \vv b \end{bmatrix}$ has a pivot in every row.
\item $\begin{bmatrix} \vv a_1 & \vv a_2 & \dots & \vv a_n & \vv b \end{bmatrix}$ has a pivot in its last (rightmost) column.
\item $\vv b$ is an image under $T$.
\item $\vv b \in \ran(T)$.
\item $\vv c \in \ran(T)$ for any choice of $\vv c$.
\item $\ran(T) = \cod(T)$.
\item $\vv a_1, \vv a_2, \dots, \vv a_n$ span $\R^m$.
\item $\vv a_1, \vv a_2, \dots, \vv a_n$ are linearly independent.
\item $\vv a_1, \vv a_2, \dots, \vv a_n, \vv b$ are linearly independent.
\item $\vv a_1, \vv a_2, \dots, \vv a_n, \vv c$ are linearly dependent for any choice of $\mathbf c$.
\end{enumerate}
\pagebreak
For any pair (i) and (j) you should be able to determine which of the following hold.
\begin{itemize}
\item if (i) is true than (j) must be true.
\item if (i) is true then (j) must be false.
\item if (i) is true then (j) may or may not be true.
\end{itemize}
For example, if (1) is true then (6) must be true. Also, if (8) is true then (18) must be false. But if (3) is true, then (7) may or may not be true.

\textbf{Suppose (5) is true. For each statement, determine if it must be true, must be false, or may or may not be true.}
(The remaining cases you will have to do on your own, this may be a good way to study the material.)

\medskip

\begin{solution}
\end{solution}

\pagebreak
\section{Linearly Dependent Sets}
(15 points) Write down three \textbf{nonzero} vectors $\mathbf v_1$, $\mathbf v_2$, and $\mathbf v_3$ in $\mathbb R^3$ such that
\begin{itemize}
\item $\{\mathbf v_1, \mathbf v_2, \mathbf v_3\}$ is linearly dependent and
\item $\mathbf v_1$ \textbf{cannot} be written as a linear combination of $\mathbf v_2$ and $\mathbf v_3$.
\end{itemize}
Your solution should be given in the following form:
\begin{displaymath}
  \mathbf v_1 =
  \begin{bmatrix}
    * \\ * \\ *
  \end{bmatrix}
  \qquad
  \mathbf v_2 =
  \begin{bmatrix}
    * \\ * \\ *
  \end{bmatrix}
  \qquad
  \mathbf v_3 =
  \begin{bmatrix}
    * \\ * \\ *
  \end{bmatrix}
\end{displaymath}
where $*$ represents an arbitrary entry.

\begin{solution}
\end{solution}

\pagebreak
\section{3D Matrix Transformations}

\begin{enumerate}
\item Find the $3 \times 3$ matrix which implements the linear transformation which rotates vectors around the $x_3$-axis.
\item Find the $3 \times 3$ matrix which implements the linear transformation which rotates vectors around the line generated by the span of the vector $\begin{bmatrix} 1 \\ 1 \\ 0\end{bmatrix}$ by 180 degrees.
\end{enumerate}

\begin{solution}
\end{solution}
\end{document}
